---
title: "MSR Final Assignment"
author: "Audrya Kerenhappukh | Cihan Acilov | Anuruddha Punam Kristhombuge"
date: "11/8/2021"
output:
  html_notebook:
    code_folding: "none"
    toc: yes
    toc_float:
      collapsed: false
      smooth_scroll: false
      
  html_document:
    theme: united
    highlight: tango
    toc: yes
    toc_float:
      collapsed: false
      smooth_scroll: false
    df_print: paged
---

```{r Setup, include=FALSE}
library(readxl)
library(tidyr)
library(tidyverse)
library(dplyr)
library(sparkline)
library(DT)
library(ggplot2)
library(e1071)
library(ggpubr)
```

## Data Preparation 
Import data set and select only limited range 
```{r}
RawDF <- read_excel("HAN_Stat & Math Foundation-Assignment _rel 02.xlsx", 
                    range = "B11:H22")
print(RawDF)

```
```{r}


CleanDF <- RawDF %>% gather(key=Courses, value=Grade, -Students)

#Example of dataset for a student
CleanDF %>% filter(Students=='a') %>% print()

```

## Data Understanding 
### Table with Trend Graphics
Use Sparkline package to create a sparkline HTML table as following 
```{r}
#Create HTML table for Histogram
course_histogram_data <- CleanDF %>%
  group_by(Courses) %>%
  summarise(
    Histogram = spk_chr(
      Grade, type ="bar",
      chartRangeMin = 100, chartRangeMax = max(Grade)
    )
  )

#Create HTML table for Sparkline
course_sparkline_data <- CleanDF %>%
  group_by(Courses) %>%
  summarise(
    Sparkline = spk_chr(
      Grade, type ="line",
      chartRangeMin = 100, chartRangeMax = max(Grade)
    )
  )

#Example of data frame:
head(course_histogram_data,2)
```
Merge the sparkline/histogram table with the current dataset. 
```{r}

#As dataset above are showing courses number as one column, we need to transpose the table so it becomes column names 
trend.rows <- left_join(course_histogram_data, course_sparkline_data) %>% t() %>% as.data.frame()

#Promote the first row as headline
colnames(trend.rows) <- trend.rows[1,]
trend.rows = trend.rows[-1, ]


trend.rows <- tibble::rownames_to_column(trend.rows, "Students")

head(trend.rows, 2)
```

Append the trend.rows with the original dataset. It is important to ensure that the column names of both data frame are the same. 
```{r}
CompleteDF <- rbind(RawDF,trend.rows)
```

Create grade trend for each students (each row)
```{r}           

#Create HTML table for Histogram
grades_histogram_data <- CleanDF %>%
  group_by(Students) %>%
  summarize(
    Histogram = spk_chr(
      Grade, type ="bar",
      chartRangeMin = 100, chartRangeMax = max(Grade)
    )
  )

# Merge the histogram with the complete data frame 
CompleteDF <- left_join(CompleteDF, grades_histogram_data, by = "Students")

#Create HTML table for Sparklines 
grades_sparkline_data <- CleanDF %>%
  group_by(Students) %>%
  summarize(
    Sparkline = spk_chr(
      Grade, type ="line",
      chartRangeMin = 100, chartRangeMax = max(Grade)
    )
  )

# Merge the histogram with the complete data frame 
CompleteDF <- left_join(CompleteDF, grades_sparkline_data)
```

Using DT library to create a HTML table with visuals:
```{r}
complete.table <- datatable(CompleteDF,  rownames = FALSE, escape = FALSE, options = list(dom = 't', paging = FALSE, ordering =  FALSE, fnDrawCallback = htmlwidgets::JS(
  '
function(){
  HTMLWidgets.staticRender();
}
'
)
)) %>%
  spk_add_deps()  

complete.table

```

### Histogram 
#### Histogram 1 (Absolute Number)

Create a Frequency and Cumulative Table based on the unpivoted column.
```{r}
grade.count <- as.data.frame(table(CleanDF$Grade)) 
names(grade.count)[names(grade.count) == "Var1"] <- "Grade"


freq.table <- transform(grade.count, Cumulative = cumsum(Freq))

freq.table
```

```{r warning=FALSE}

ratio <- max(freq.table$Freq) / max(freq.table$Cumulative) #Create a ratio variable for the Y axis limit 

freq.graph <- ggplot(freq.table, aes(x = Grade)) + 
    # Create a bargraph for the absolute frequency
  geom_bar(aes(y=Freq, fill="Frequency"), size = 1, stat="identity") + 
     # Create a line graph for Cumulative Count
   geom_line(aes(y = Cumulative*ratio, colour = "Cumulative"), group=1, size=1) +
    # Create labels 
   labs(title = "Overall Distribution of Grades", x = "Grades", y = "Frequency" ) + 
    # Scale the secondary values 
   scale_y_continuous(breaks=seq(0,max(freq.table$Freq),by = 2),
                      sec.axis=sec_axis(~./ratio,name="Cumulative", breaks=seq(0,max(freq.table$Cumulative), by = 10))) +
    #Set color and legend setings 
   scale_colour_manual("Cumulative", values=c("Cumulative" = "tan1"))+
   scale_fill_manual("Frequency",values="grey")+
    #Set plot design 
   theme(plot.title = element_text(hjust = 0.5), #Title in the center
         #Legend layout:
         legend.key=element_blank(),
         legend.title=element_blank(),
         legend.box="horizontal", 
         legend.position = "top")

freq.graph
 
```

Frequency show that how many times appearing on the grades and culminative frequency distribution in the sum of the class.  Grade 4 students performance very high and grage 10 below the average. 


#### Histogram 2 (Relative Number)
Create a proportion table for the relative values based on the freq.table
```{r}
prop.table <- as.data.frame(cbind(Grade = freq.table$Grade,
                                  Freq=prop.table(freq.table$Freq),
                                  Cumulative = cumsum(prop.table(freq.table$Freq))))
prop.table$Grade <- factor(prop.table$Grade, levels = c(1:10))

prop.table
```

```{r}

ratio <- max(prop.table$Freq) / max(prop.table$Cumulative) 


 relative.graph <- ggplot(prop.table, aes(x = Grade)) + 
 #Create a bargraph
  geom_bar(aes(y=Freq, fill="Frequency"), size = 1, stat="identity") + 
   #Create a line graph for cumulative 
   geom_line(aes(y = Cumulative*ratio, colour = "Cumulative"), group=1, size=1) + 
    #Add labels 
   labs(title = "Overall Distribution of Grades", x = "Grades", y = "Frequency" ) + 
    #Adjust grid major lines and secondary axis 
   scale_y_continuous(breaks=seq(0,max(prop.table$Freq),by = 0.02), 
                                 labels = scales::percent,
                                 sec.axis=sec_axis(~./ratio,name="Cumulative",
                                          breaks=seq(0,max(prop.table$Cumulative), by = 0.1), labels = scales::percent)) +
    # Add Legend and adjust legend and title positions
   scale_colour_manual("Cumulative", values=c("Cumulative" = "tan1"))+
   scale_fill_manual("Frequency",values="grey")+
   theme(plot.title = element_text(hjust = 0.5),
         legend.key=element_blank(),
         legend.title=element_blank(),
         legend.box="horizontal", 
         legend.position = "top")
 
 relative.graph
```
Frequency show that percentage of how many times appearing on the grades and culminative percentage of frequency distribution in the sum of the class.  Grade 4 students performance very high and grage 10 below the average.


### Graphics by Students
Create a table with all the calculations
```{r}

#Because current Stdev function in R is Stdev for Sample, we need to create a function 
sd.p <- function(x){sd(x)*sqrt((length(x)-1)/length(x))}

#Create a table for everyting
stat.table <- cbind(Students = RawDF$Students, 
                    Sum = rowSums(RawDF[,-1]),
                    Mean = round(rowMeans(RawDF[,-1]),2),
                    Stdev = round(apply(RawDF[,-1],1,sd.p),2),
                    Skewness = round(apply(RawDF[,-1],1,skewness,na.rm= TRUE),2),
                    Kurtosis = round(apply(RawDF[,-1],1,kurtosis,na.rm= TRUE),2)) %>% as.data.frame

#Because the results of function above showing numeric values, we need to transform them to numeric values
stat.table <- as.data.frame(cbind(Students = stat.table$Students, as.data.frame( lapply(stat.table[,-1], as.numeric))))

#Add the table for Mean of Means 
stat.table <- transform(stat.table, MeanOfMeans = mean(stat.table$Mean), MeanStd = mean(stat.table$Stdev), MeanSkew = mean(stat.table$Skewness), MeanKurtosis = mean(stat.table$Kurtosis))

head(stat.table)
```

##### Means Grade by Student
```{r}

#A variable for Y axes limit 
y.grid <- seq(min(RawDF[,-1]), max(RawDF[,-1]), by = 1)

student.mean <- ggplot(stat.table,aes(x = Students)) + 
   geom_line(aes(y = Mean, colour = "Mean"), size = 1 , group = 1) + 
   geom_point(aes(y = Mean,  colour = "Mean")) +
   geom_line(aes (y = MeanOfMeans , colour = "Mean of Means"), group = 1, size = 1, linetype = "dashed") +
   scale_y_continuous(limits = c(min(RawDF[,-1]), max(RawDF[,-1])),  breaks = y.grid )+ 
   labs(title = "Means Grade by Student", x ="Students", y="") + 
    scale_colour_manual(values=c("Mean" = "blue", "Mean of Means" = "red")) +
    theme(plot.title = element_text(hjust = 0.5),
         legend.key=element_blank(),
         legend.title=element_blank(),
         legend.box="horizontal", 
         legend.position = "top")

student.mean

```
##### Standard Deviation of Grades by Student
```{r}
student.stdev <- ggplot(stat.table,aes(x = Students)) + 
   geom_line(aes(y = Stdev, colour = "Standard Deviation"), size = 1 , group = 1) + 
   geom_point(aes(y = Stdev,  colour = "Standard Deviation")) +
   geom_line(aes (y = MeanStd , colour = "Mean of Std"), group = 1, size = 1, linetype = "dashed") +
   scale_y_continuous(limits = c(0,3.5),  breaks = seq(0,3.5,by =0.5))+ 
   labs(title = "Standard Deviation of Grades by Student", x ="Students", y="") + 
   scale_colour_manual(values=c("Standard Deviation" = "blue", "Mean of Std" = "red")) +
    theme(plot.title = element_text(hjust = 0.5),
         legend.key=element_blank(),
         legend.title=element_blank(),
         legend.box="horizontal", 
         legend.position = "top")
student.stdev
```
##### Skewness of Grades by Student
```{r}
student.skewness <- ggplot(stat.table,aes(x = Students)) + 
   geom_line(aes(y = Skewness, colour = "Skewness"), size = 1 , group = 1) + 
   geom_point(aes(y = Skewness,  colour = "Skewness")) +
   geom_line(aes (y = MeanSkew , colour = "Mean of Skewness"), group = 1, size = 1, linetype = "dashed") +
   scale_y_continuous(limits = c(min(stat.table$Skewness) - 0.5, max (stat.table$Skewness) + 0.5),  breaks = seq(min(stat.table$Skewness) - 0.5, max (stat.table$Skewness) + 0.5,by =0.2))+ 
   labs(title = "Skewness of Grades by Student", x ="Students", y="") + 
   scale_colour_manual(values=c("Skewness" = "blue", "Mean of Skewness" = "red")) +
    theme(plot.title = element_text(hjust = 0.5),
         legend.key=element_blank(),
         legend.title=element_blank(),
         legend.box="horizontal", 
         legend.position = "top")
student.skewness
```
##### Kurtosis of Grades by Student
```{r}
student.kurtosis <- ggplot(stat.table,aes(x = Students)) + 
   geom_line(aes(y = Kurtosis, colour = "Kurtosis"), size = 1 , group = 1) + 
   geom_point(aes(y = Kurtosis,  colour = "Kurtosis")) +
   geom_line(aes (y = MeanKurtosis , colour = "Mean of Kurtosis"), group = 1, size = 1, linetype = "dashed") +
   scale_y_continuous(limits = c(min(stat.table$Kurtosis)-0.5, max(stat.table$Kurtosis)+0.5),  breaks = seq( min(stat.table$Kurtosis)-0.5, max(stat.table$Kurtosis)+0.5, by =0.5))+ 
   labs(title = "Kurtosis of Grades by Student", x ="Students", y="") + 
   scale_colour_manual(values=c("Kurtosis" = "blue", "Mean of Kurtosis" = "red")) +
    theme(plot.title = element_text(hjust = 0.5),
         legend.key=element_blank(),
         legend.title=element_blank(),
         legend.box="horizontal", 
         legend.position = "top")
student.kurtosis
```

### Graphics by Course
Create a transpose table for seeing the Course average
```{r}
courseDF <- RawDF %>% t() %>% as.data.frame()

#Change the first row as column names 
colnames(courseDF) <- courseDF[1,]
courseDF <- courseDF[-1, ]

#Change row names as a new column for Course
courseDF <- tibble::rownames_to_column(courseDF, "Course")

#Change all course columns to Numeric
courseDF <- cbind("Course"= courseDF[,1], as.data.frame(lapply(courseDF[,-1], as.numeric)))
```


```{r}
sd.p <- function(x){sd(x)*sqrt((length(x)-1)/length(x))}
   
course.table <- cbind(Course = courseDF$Course, 
                    Sum = rowSums(courseDF[,-1]),
                    Mean = round(rowMeans(courseDF[,-1]),2),
                    Stdev = round(apply(courseDF[,-1],1,sd.p),2),
                    Skewness = round(apply(courseDF[,-1],1,skewness,na.rm= TRUE),2),
                    Kurtosis = round(apply(courseDF[,-1],1,kurtosis,na.rm= TRUE),2)) %>% as.data.frame

course.table <- as.data.frame(cbind(Course = course.table$Course, as.data.frame( lapply(course.table[,-1], as.numeric))))

# Add Mean of Means Table 
course.table <- transform(course.table, MeanOfMeans = mean(course.table$Mean), MeanStd = mean(course.table$Stdev), MeanSkew = mean(course.table$Skewness), MeanKurtosis = mean(course.table$Kurtosis))
```


##### Means Grade by Course
```{r echo=TRUE}

y.grid <- seq(min(courseDF[,-1]), max(courseDF[,-1]), by = 1)

course.mean <- ggplot(course.table,aes(x = Course)) + 
   geom_line(aes(y = Mean, colour = "Mean"), size = 1 , group = 1) + 
   geom_point(aes(y = Mean,  colour = "Mean")) +
   geom_line(aes (y = MeanOfMeans , colour = "Mean of Means"), group = 1, size = 1, linetype = "dashed") +
   scale_y_continuous(limits = c(min(courseDF[,-1]), max(courseDF[,-1])),  breaks = y.grid )+ 
   labs(title = "Means Grade by Course", x ="Course", y="") + 
    scale_colour_manual(values=c("Mean" = "blue", "Mean of Means" = "red")) +
    theme(plot.title = element_text(hjust = 0.5),
         legend.key=element_blank(),
         legend.title=element_blank(),
         legend.box="horizontal", 
         legend.position = "top")
course.mean
```

##### Standard Deviation of Grades by Course
```{r echo=TRUE}
course.stdev <- ggplot(course.table,aes(x = Course)) + 
   geom_line(aes(y = Stdev, colour = "Standard Deviation"), size = 1 , group = 1) + 
   geom_point(aes(y = Stdev,  colour = "Standard Deviation")) +
   geom_line(aes (y = MeanStd , colour = "Mean of Std"), group = 1, size = 1, linetype = "dashed") +
   scale_y_continuous(limits = c(0,3.5),  breaks = seq(0,3.5,by =0.5))+ 
   labs(title = "Standard Deviation of Grades by Course", x ="Course", y="") + 
   scale_colour_manual(values=c("Standard Deviation" = "blue", "Mean of Std" = "red")) +
    theme(plot.title = element_text(hjust = 0.5),
         legend.key=element_blank(),
         legend.title=element_blank(),
         legend.box="horizontal", 
         legend.position = "top")
course.stdev
```

##### Skewness of Grades by Course
```{r echo=TRUE}
course.skewness <- ggplot(course.table,aes(x = Course)) + 
   geom_line(aes(y = Skewness, colour = "Skewness"), size = 1 , group = 1) + 
   geom_point(aes(y = Skewness,  colour = "Skewness")) +
   geom_line(aes (y = MeanSkew , colour = "Mean of Skewness"), group = 1, size = 1, linetype = "dashed") +
   scale_y_continuous(limits = c(min(course.table$Skewness) - 0.5, max (stat.table$Skewness) + 0.5),  breaks = seq(min(course.table$Skewness) - 0.5, max (course.table$Skewness) + 0.5,by =0.2))+ 
   labs(title = "Skewness of Grades by Course", x ="Courses", y="") + 
   scale_colour_manual(values=c("Skewness" = "blue", "Mean of Skewness" = "red")) +
    theme(plot.title = element_text(hjust = 0.5),
         legend.key=element_blank(),
         legend.title=element_blank(),
         legend.box="horizontal", 
         legend.position = "top")
course.skewness
```

##### Kurtosis of Grades by Course
```{r echo=TRUE}
course.kurtosis <- ggplot(course.table,aes(x = Course)) + 
   geom_line(aes(y = Kurtosis, colour = "Kurtosis"), size = 1 , group = 1) + 
   geom_point(aes(y = Kurtosis,  colour = "Kurtosis")) +
   geom_line(aes (y = MeanKurtosis , colour = "Mean of Kurtosis"), group = 1, size = 1, linetype = "dashed") +
   scale_y_continuous(limits = c(min(stat.table$Kurtosis)-0.5, max(stat.table$Kurtosis)+0.5),  breaks = seq(min(stat.table$Kurtosis)-0.5, max(stat.table$Kurtosis)+0.5, by =0.5))+ 
   labs(title = "Kurtosis of Grades by Course", x ="Courses", y="") + 
   scale_colour_manual(values=c("Kurtosis" = "blue", "Mean of Kurtosis" = "red")) +
    theme(plot.title = element_text(hjust = 0.5),
         legend.key=element_blank(),
         legend.title=element_blank(),
         legend.box="horizontal", 
         legend.position = "top")
course.kurtosis
```
## Summary
```{r warning=FALSE}
#Print HTML table
complete.table
```

```{r}
#Print Dstribution Graphs
ggarrange(freq.graph, relative.graph, nrow=1, ncol=2)
```
The graph is showing the distribution of grades in numbers. The line value is showing the cumulative count of grades, which is equel to 66 (11 students, 6 course = 66 grades)

The graph is showing the distribution of grades in percentile. The line value is showing the cumulative count of grades in percentile. 


```{r warning=FALSE}
#Summary of all Students Graphs
ggarrange(student.mean, student.stdev, student.skewness,student.kurtosis)
```
Means grade by student
Blue line shows the mean of student grades. Red line shows the mean of mean. Student C perform the highest and the student D & J performed the lowest. 
Standard deviation of grades by student
Blue line shows the standard deviation of student grades.Red line shows the mean of students. Student C & K has lowest standard deviations and the student A has the highest.
Skewness of Grades by student
Blue line shows the skewness of student grades. Red line shows the mean of skewness. Mean of skewness is below 0. 
Kurtosis of grades by student
Blue line shows the kurtosis of student grades. Red line shows the mean of kurtosis. Mean of kurtosis is below 0.  





```{r}
#Summary of all Courses Graphs
ggarrange(course.mean, course.stdev, course.skewness,course.kurtosis)
```
Means grade by course
Blue line shows the mean grade of courses. Red line shows the mean of mean. course 6 perform the highest and the course 4 performed the lowest. 
Standard deviation of grades by course
Blue line shows the standard deviation of courses. Red line shows the mean of student. course 6 has the lowest standard deviation and the course 1 has the highest. 
Skewness of Grades by course
Blue line shows the skewness of courses. Red line shows the mean of skewness. course 4 has the highest skewness and the course 5 has the lowest. 
Kurtosis of grades by course
Blue line shows the kurtosis of courses. Red line shows the mean of kurtosis. course 4 has the highest kurtosis and the course 3 has the lowest. 
